---
title: "R Notebook"
output: html_notebook
---

#Before we start

Running libaries 

```{r message=FALSE}
library(tidyverse)

library(lubridate)

library(magrittr)
```


```{r}
pokemon= read_csv("https://sds-aau.github.io/SDS-master/00_data/pokemon.csv")
```

#Tasks

**Give a brief overview of data, what variables are there, how are the variables scaled and variation of the data columns.**

**Execute a PCA analysis on all numerical variables in the dataset. Hint: Don’t forget to scale them first. Use 4 components. What is the cumulative explained variance ratio? Hint: I am not sure this terminology and code was introduced during class, but try and look into cumulative explained variance and sklearn(package) and see if you can figure out the code needed.**
****

**Use a different dimensionality reduction method (eg. UMAP/NMF) – do the findings differ?**


**Perform a cluster analysis (KMeans) on all numerical variables (scaled & before PCA). Pick a realistic number of clusters (up to you where the large clusters remain mostly stable).**


**Visualize the first 2 principal components and color the datapoints by cluster.**

**Inspect the distribution of the variable Type1 across clusters. Does the algorithm separate the different types of pokemon?**


**Perform a cluster analysis on all numerical variables scaled and AFTER dimensionality reduction and visualize the first 2 principal components.**

**Again, inspect the distribution of the variable “Type 1” across clusters, does it differ from the distribution before dimensionality reduction?**

